Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-02-12 bbbq_article.tex
Remove obsolete text, add braces around refs

It is helpful to be able to predict which peptides are immunogenic,
in, among others, vaccine development. 
Already for a decade, synthesized peptide fragments are used
to experimentally determine which peptides
are immunogenic.
This approach, however, is tedious and costly.
Due to this, software was written that allows for silico 
predictions, which are reliable in 
practice \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.

@@

It is unknown if the presentation of hydrophobic/TMH residues
is evolutionary selected for.
One possible explanation is that TMHs have a reduced variability,
imposed by the functional requirement of being able to span a lipid bilayer.
Due to this, the TMHs that pathogens possess, 
have a lower chance to develop an escape mutation,
as many mutations will result in a dysfunctional TMH.

@@

\paragraph{Data set for evolutionary conservation of TMHs}

To detect the evolutionary conservation of TMHs,
we use the 49790 (as of 2020-06-22) fully sequenced SARS-CoV-2 proteomes, 
as provided by GISAID (\cite{shu2017gisaid}).

@@

To determine if epitopes derived from TMHs are presented at all,
we started from epitopes sequenced from elution studies
for MHC-I \cite{schellens2015comprehensive} 
and MHC-II \cite{bergseng2015different}.
For each epitope, the human representative reference proteome
was searched for its origin.
We kept only the epitopes of which were found exactly once
on the proteome.
We predicted the topology of the proteome
and counted how often the epitopes overlapped
with a predicted TMH.
In this analysis, both TMHMM and 
PureseqTM (see section 'Prediction software used') 
were used to predict the topology.
The full analysis can be found
at \url{https://github.com/richelbilderbeek/bbbq_article_issue_157}.

@@

To determine the percentages of MHC-I and MHC-II epitopes overlap
with TMHs, we used mostly the same analysis as described in \cite{bianchi2017}.
To summarize: from a proteome all possible 9-mers are derived. For each
of these peptides, it is measured if it has been part of a 
TMH (at least overlapping 1 residue), 
as well as if the peptide is a binder to an MHC-I haplotype.
For MHC-II, 14-mers are used, as these are the most occurring
epitope length \cite{bergseng2015different}.

Because the percentage of TMH residues is known, a binomial test can be
used to determine if TMH-derived peptides are as likely to be a binder
as expected by chance (the null hypothesis), or that TMHs are indeed an
overrepresented source of epitopes (as found by \cite{bianchi2017}).

@@

The definition of a binder differs from \cite{bianchi2017}:
in this study a peptide is called a binder if, within a haplotype, 
any of its 9-mer peptides have an IC50 value in the lowest 2\% of 
the epitopes within a proteome (see table \ref{table:ic50s_per_haplotype} for values), 
whereas the original study defined
a binder as having an IC50 in the lowest 2\% of the peptides within a protein.
% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17

The 13 MHC-I haplotypes used in this study are the same as 
the ones in \cite{bianchi2017}.
The MHC-II haplotypes used additionally are selected 
to occur with a phenotypic frequency of at least 14\% in
the human population (\cite{greenbaum2011functional},
see \ref{table:mhc2_haplotypes} for coverage),
resulting in 21 haplotypes.
When using an MHC-II haplotype, instead of using 9-mers, 14-mers are
used, as these are the most common MHC-II epitope size,
as found in the elution study (\cite{bergseng2015different}).

@@

For MHC-I, there are multiple computational tools developed 
to predict epitopes. 
According to \cite{lundegaard2011prediction}, in 2011,
from a set of multiple tools, 
NetMHCcons (\cite{karosiene2012netmhccons}) gave the best predictions.
A tool developed later is \verb;epitope-prediction; (\cite{bianchi2017}),
which uses a stabilized matrix method (\cite{kim2009derivation}).
The most recent and promising FOSS tool available now appears
to be MHCnuggets (\cite{shao2020high}), which can do both MHC-I 
and MHC-II prediction.

>>>>

It is helpful to be able to predict which peptides are immunogenic,
in, among others, vaccine development. 
Already for a decade, synthesized peptide fragments are used
to experimentally determine which peptides
are immunogenic.
This approach, however, is tedious and costly.
Due to this, software was written that allows for silico 
predictions, which are reliable in 
practice (\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}).

@@

It is unknown if the presentation of hydrophobic/TMH residues
is evolutionary selected for.
One reason to expect a reduced variability (and hence 
evolutionary conservation) in TMHs, is
that these are restricted in their evolution
by the functional requirement to span a lipid bilayer.
Due to this, the TMHs that pathogens possess, 
have a lower chance to develop an escape mutation,
as many mutations will result in a dysfunctional TMH.

@@

@@

To determine if epitopes derived from TMHs are presented at all,
we started from epitopes sequenced from elution studies
for MHC-I (\cite{schellens2015comprehensive}) 
and MHC-II (\cite{bergseng2015different}).
For each epitope, the human representative reference proteome
was searched for its origin.
We kept only the epitopes of which were found exactly once
on the proteome.
We predicted the topology of the proteome
and counted how often the epitopes overlapped
with a predicted TMH.
In this analysis, both TMHMM and 
PureseqTM (see section 'Prediction software used') 
were used to predict the topology.
The full analysis can be found
at \url{https://github.com/richelbilderbeek/bbbq_article_issue_157}.

@@

To determine the percentages of MHC-I and MHC-II epitopes overlap
with TMHs, we used mostly the same analysis as described in \cite{bianchi2017}.
To summarize: from a proteome all possible 9-mers are derived. For each
of these peptides, it is measured if it has been part of a 
TMH (at least overlapping 1 residue), 
as well as if the peptide is a binder to an MHC-I haplotype.
For MHC-II, 14-mers are used, as these are the most frequently occurring
epitope length (\cite{bergseng2015different}).

@@

The definition of a binder differs from \cite{bianchi2017}:
in this study a peptide is called a binder if, within a haplotype, 
any of its 9-mer peptides have an IC50 value in the lowest 2\% of 
the epitopes within a 
\emph{proteome} (see tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
for values), whereas the original study defined
a binder as having an IC50 in the lowest 2\% 
of the peptides within a \emph{protein}.
% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17

The 13 MHC-I haplotypes used in this study are the same as 
the ones in \cite{bianchi2017}.
The MHC-II haplotypes used additionally are selected 
to occur with a phenotypic frequency of at least 14\% in
the human population (\cite{greenbaum2011functional}),
resulting in 21 haplotypes.
When using an MHC-II haplotype, instead of using 9-mers, 14-mers are
used, as these are the most common MHC-II epitope size,
as found in the elution study (\cite{bergseng2015different}).

@@

For MHC-I, there are multiple computational tools developed 
to predict epitopes. 
According to \cite{lundegaard2011prediction}, at that time,
NetMHCcons (\cite{karosiene2012netmhccons}) gave the best predictions.
A tool developed later is \verb;epitope-prediction; (\cite{bianchi2017}),
which uses a stabilized matrix method (\cite{kim2009derivation}).
The most recent and promising FOSS tool available now appears
to be MHCnuggets (\cite{shao2020high}), which can do both MHC-I 
and MHC-II prediction.
